---
title: "R Notebook"
output: html_notebook
---


# Source and documentation

* [Deepankar Basu academic profile](https://people.umass.edu/dbasu/data.html)
* [EPWT Documentation](https://people.umass.edu/dbasu/Data/Documentation%20EPWT%207.0.pdf)
* [Pagina web oficial PWT](https://www.rug.nl/ggdc/productivity/pwt/)
* [The Next Generation of the Penn World Table](https://www.rug.nl/ggdc/docs/the_next_generation_of_the_penn_world_table.pdf)(paper que detalla la metodología y algunos resultados)
* [CRAN documentation](https://cran.r-project.org/web/packages/pwt10/pwt10.pdf)



Selected variables
* `rnat2017`: Net profit rate measured at constant 2017 national prices.
* `rnatcur` _ It is the net profit rate measured at current national prices.
  * It is computed as: [(1 – LabShare)* rhocur] – delta]
  * I supposed rhocur is  rhonatcur _ It is the productivity of capital (output-capital ratio) measured at current national prices. It is the ratio between the real GDP at current national prices and the capital stock at current national prices.
  * It is their preferred measure of profit rate
  
  rnatcur=[(1-Labshare)*XGDPnatcur-(delta*Knatcur)]/Knatcur
  

* `rppp2017`: Net profit rate measured at 2017 PPPs.

prices. It is the ratio between the real GDP at current national prices and the capital stock at
current national prices.



tg = [(1-labs)-delta]* rhotcur


# Data


```{r}
require(tidyverse)
require(readxl)
require(reshape2)
require(ggplot2)
```


```{r}
preproc <- function(data, country, unidades=1e3){
  data <- data  %>% 
  mutate(iso3 = country,
         fuente='Propia',
         LabShare = rem/pib,
         delta = ckf/stock,
         pib = pib* unidades,
         rem=rem*unidades, 
         pv=pv*unidades,
         stock=stock*unidades ,
         ckf=ckf*unidades
         
         ) %>% 
  select(var_sel_pict) %>% 
  melt(id.vars=c('anio', 'iso3', 'fuente'))
  return(data) 
  
}
```


```{r}
# selected countries and vars
paises_sel <- c("ECU", "URY","CHL", "VEN"
              # ,"ARG","BRA", "BOL",
              #   "PER","COL","PRY","MEX"
              )

var_sel_pict <- c("anio", 'iso3', 'fuente',
                  "pib", "rem", "pv" ,
                  "stock" ,  "ckf",
                  'tg', 'LabShare',
                  'delta'
                  # , "kta" , "kta_t_1",   "kcca"
                    
                  )

# epwt data
epwt <- read_excel( '../../data/pwt/EPWT 7.0 FV.xlsx',
                   sheet = 'EPWT7.0') %>% 
  filter(Countrycode %in% paises_sel ) %>%
  filter(!(Countrycode == 'VEN' & Year>=2018))  %>%
  select('anio' = Year, 'iso3'= Countrycode, # 'pais' = Country,
          "LabShare",
          'pib'="XGDPnatcur",
          "rhonatcur" ,
          'delta', 
          'stock'= "Knatcur", "knat2017",
          'rnat2017', 'tg' = 'rnatcur', 'rppp2017') %>% 
  mutate(rem=LabShare*pib,
         pv = ((1 - LabShare) * rhonatcur) * stock ,
         fuente= 'EPWT',
         ckf = delta * stock
         ) %>% 
  select(var_sel_pict) %>% 
    melt(id.vars=c('anio', 'iso3', 'fuente'))

  
# epwt_vars <- read_excel('../../data/pwt/EPWT 7.0 FV.xlsx',
#                    sheet = 1)

# pict data
df_uy <- read_excel('../../data/pict/base_total_economía_uruguay output.xlsx',
                    sheet = 2) %>% 
  preproc(., 'URY')



df_ec <- read_excel('../../data/pict/base_total_economía_ecuador.xlsx',
                    sheet = 2) %>% 
  preproc(., 'ECU') 


df_cl <- read_excel('../../data/pict/base_total_economía_chile.xlsx',
                    sheet = 2) %>% 
  preproc(., 'CHL', 1e6) 

df_vz <- read_excel('../../data/pict/base_total_economía_venezuela.xlsx',
                    sheet = 2) %>% 
  preproc(., 'VEN') 


# bind data
df_bind <- bind_rows(df_ec, df_uy) %>% 
  bind_rows(df_cl) %>% 
  bind_rows(df_vz)

data <-  df_bind %>% 
  bind_rows(epwt)

# join data
df_join <- epwt %>% 
  select(-fuente) %>% 
  rename(value_epwt = value) %>% 
  left_join(df_bind %>% select(-fuente), by = join_by(anio, iso3, variable)) %>% 
  mutate(brecha = value/value_epwt-1)

head(data)
```

```{r}
unique(data$variable)
```


#Visualization
## Valores
```{r}

# Extract unique variables from the 'variable' column
variables <- unique(data$variable)

# Loop through each variable and create a line plot
for (var in variables) {
  # Eje y
  if(!(var %in% c('delta', 'tg', 'LabShare')) ){
    y_lab =  "Moneda nacional a precios corrientes"
  }else{
    y_lab =  ""
    }
  
  # Subset data for the current variable
  subset_data <- data %>%
     filter(variable == var) #%>% 
    # filter(fuente=='Propia')
    # filter(fuente=='EPWT')
  
  # Create a ggplot object
  p <- ggplot(subset_data, aes(x = anio, y = value, color = fuente)) +
    geom_line(alpha = 0.6) +
    # geom_point()+
    labs(title = var, x = "Año", y = y_lab) +
    facet_wrap(~ iso3, scales='free',
               ncol=2) + # Facet by country
    scale_color_discrete(name = "Fuente") # Optional: change legend title
  
  # Save
  ggsave(paste0('../../results/pict/comparacion_epwt/', var, '.png' ), plot = p)
  
  # Print the plot
  print(p)
}

```

## Brecha

```{r}

# Extract unique variables from the 'variable' column
variables <- unique(df_join$variable)

# Loop through each variable and create a line plot
for (var in variables) {
  # Subset data for the current variable
  subset_data <- df_join %>%
     filter(variable == var) #%>% 
    # filter(fuente=='Propia')
    # filter(fuente=='EPWT')
  
  # Create a ggplot object
  p <- ggplot(subset_data, aes(x = anio, y = brecha*100, fill = iso3)) +
    # geom_line(alpha = 0.6) +
    geom_col()+
    labs(title = var, x = "Year", y = "%") +
    facet_wrap(~ iso3,# scales='free',
               ncol=2) + # Facet by country
    scale_color_discrete(name = "Source") # Optional: change legend title
  
  # Save
  ggsave(paste0('../../results/pict/comparacion_epwt/', var, '_brecha_misma_escala.png' ), plot = p)
  
  # Print the plot
  print(p)
}

```



